---
title: "HR Analytics - Summative Assessment"
author: your_id_here
output: 
  html_document: 
    theme: flat
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---
\pagebreak
```{r setup, include=FALSE}
rm(list=ls())
gc()

library(tidyverse)
 library(corrplot)     
 library(randomForest)
 library(knitr)
 library(kableExtra)
 library(pbapply)
 library(Matrix)
 library(factoextra)
 library(xgboost)
 library(fastDummies)
 library(ggrepel)
 library(glmnet)
  library(janitor)
library(ROSE)
library(rpart)
```

## Read the data

Be sure to ask questions here:
<https://docs.google.com/document/d/128siy6JNSyG1r56QloVX4B_3-l501oSOhAgn6YS2rHc/edit?usp=sharing>

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv')
glimpse(hr)
summary(hr)
```

## Duplicate check

Check for and remove duplicates: This chunk finds that there are many people who have the exact same details and it's duplicates more than 3 times.

```{r}
hr %>% 
  group_by_all() %>% 
  summarize(n = n()) %>% 
  group_by(left) %>% 
  summarise(mean_duplicates = mean(n))

```

```{R Maximum duplicate check}
hr %>% 
  group_by_all() %>% 
  summarize(n = n()) %>% 
  group_by(left) %>% 
  summarise(max_duplicates = max(n))
```

```{r}
d <- hr %>% 
  group_by_all() %>% 
  summarize(dupe_count = n()) %>% 
  ungroup() %>% 
  mutate(left = ifelse(left == 1, 'Left', 'Stayed'))

ggplot(filter(d, dupe_count > 0), 
       aes(x = dupe_count, group = left, fill = left, y = ..prop..)) +
  geom_bar(position = 'dodge', alpha = 2, width = 0.5) +
  scale_x_continuous(breaks = 1:6, labels = 1:6) +
  labs(title = "Occurance of data")

```

>Base on the data, there ar  more leavers than stayers. This means the day tends to be more biase to leavers than stayers. 


Remove the duplicates
```{r}
hr_new <- unique(hr) 
```

```{R probability}
prop.table(table(hr_new$left))
```

```{R oversamping }
hr_new_oversample <- ovun.sample(left ~., data = hr_new, method = "both", p=0.5, N = 14999, seed = 1)$data
hr_new_oversample <- as.numeric(hr_new_oversample)
table(hr_new_oversample$left)
```
## Linear regression of satisfaction

```{r}
sat_lm <- lm(satisfaction_level ~ ., data = hr_new_oversample)
summary(sat_lm)
plot(sat_lm)
```

Add an ID variable
```{r}
hr_new_oversample$id <- as.character(1:nrow(hr_new_oversample))
```

## Correlation plot

```{r}
library(corrplot)
m_cor <- hr_new_oversample %>% 
  select_if(is.numeric) %>% 
  cor()
corrplot(m_cor, method = 'color', diag = F, tl.col = 'grey20', addrect = 2, order = 'hclust')
```

```{r}
reshape2::melt(m_cor) %>% 
  filter(value != 1) %>% 
  arrange(abs(value)) %>% 
  filter(Var1 == 'left')
```
>We want to see what correlated with leaving

Create dummy columns for these variables.

```{r}
library(fastDummies)
hr <- dummy_cols(hr, select_columns = c('salary', 'sales'))
```

## PCA
>Shows the dimention and varience

```{r}
library(factoextra)
library(FactoMineR)

hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  select(-sales)

hr_pca <- hr %>% 
  select_if(is.numeric) %>% 
  PCA

hr_num <- hr %>% 
  select(satisfaction_level, last_evaluation, number_project, average_montly_hours, time_spend_company, Work_accident, promotion_last_5years)

hr_pca <- PCA(hr_num)
fviz_pca_biplot(hr_pca, geom = 'point', col.ind = hr$left, axes = c(2,3))
```

```{R new}
fviz_eig(hr_pca)
```

```{R kmeans cluster plot}
hr1 <- hr %>% 
  filter(left == 1) %>% 
  select(satisfaction_level, last_evaluation)

km1 <- hr1 %>% 
  kmeans(centers = 3, nstart = 20)

hr1$cluster <- as.character(km1$cluster)

ggplot(hr1, aes(x = satisfaction_level, y = last_evaluation, color = cluster)) +
  geom_point(alpha = 0.5)

```
## Histogram of departments

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3))

hr <- hr %>% 
  mutate(sales = case_when(sales == 'sales' ~ 'Sales',
                           sales == 'technical' ~ 'Technical',
                           T ~ sales)) %>% 
  rename(department =  sales) %>% 
  mutate(department = fct_infreq(department))

ggplot(hr, aes(x = salary)) +
  geom_bar()

```

## Leavers by department

```{r}
d <- hr %>% 
  group_by(department, left) %>%
  summarize(n = n()) %>% 
  mutate(proportion = n / sum(n))

ggplot(d) +
  geom_bar(stat = 'identity', position = 'stack',
           aes(x = department, y = n, fill= left))
```
```{R newset}
d %>% 
  ungroup() %>% 
  filter(left == 1) %>% 
  mutate(department = fct_reorder(department, proportion)) %>% 
  ggplot() +
  geom_segment(x = 0, aes(xend = proportion, y = department, yend = department)) +
  geom_point(aes(y = department, x = proportion, size = n)) +
  scale_x_continuous('Proportion of Leavers')
```

## Tenure

```{r}
d <- hr %>% 
  group_by(time_spend_company) %>% 
  summarize(n = n()) %>% 
  mutate(p = n / sum(n))

ggplot(hr, aes(x = time_spend_company)) +
  geom_bar() +
  scale_x_continuous("Tenure (years)",breaks = 1:10, limits = c(1, 11)) +
  theme_minimal() +
  theme(panel.grid = element_blank())

```

## Tenure and satisfaction

```{r}
a <- hr_new_oversample %>% 
  group_by(sales, left) %>%
  summarize(n = n()) %>% 
  mutate(proportion = n / sum(n))

ggplot(a, aes(x = time_spend_company, y = satisfaction_level)) +
  pie(alpha = 0.2) +
  scale_x_continuous(breaks = 1:10) +
  theme_minimal() + theme(panel.grid = element_blank())
```

```{R pie chart}
pie <- ggplot(hr_new_oversample, aes(x = n , y= satisfaction_level,fill = factor(sales))) + 
  geom_bar(width = 1, stat = "identity") +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5))

pie + coord_polar(theta = "y", start=0)
```    
## Tenure and Department

```{r}
ggplot(hr_new_oversample, aes(x = time_spend_company y = ), fill(department)) +
  geom_point(alpha = 0.2) +
  scale_y_continuous(breaks = 1:10) +
  theme_minimal() + theme(panel.grid = element_blank())
```


```{r}
d <- hr %>% 
  group_by(department, time_spend_company)%>% 
  summarize(n = n()) %>% 
  mutate(p = n / sum(n))

ggplot(d, aes(x = time_spend_company, y = p, fill = department)) +
  geom_bar(stat = 'identity', position = 'fill') +
  scale_fill_brewer(type = 'qual', palette = 3) +
  scale_x_continuous("Tenure (years)",breaks = 1:10, limits = c(1, 11)) +
  theme_minimal() +
  theme(panel.grid = element_blank())
```



## K-means

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  select(-sales) %>% 
  mutate(id = row_number())

hr1 <- hr %>% 
  filter(left == 1) %>% 
  select(id, satisfaction_level, last_evaluation)

km1 <- hr1 %>% 
  select(-id) %>% 
  kmeans(centers = 3, nstart = 20)

hr1$cluster <- as.character(km1$cluster)

hr_clust <- left_join(hr, 
          select(hr1, id, cluster), 
          by = 'id') %>% 
  mutate(cluster = ifelse(is.na(cluster), "stayed", cluster))
  

ggplot(hr_clust, aes(x = satisfaction_level, y = last_evaluation, color = cluster)) +
  geom_point(alpha = 0.5)

ggplot(hr_clust, aes(x = number_project, y = average_montly_hours, color = cluster)) +
  geom_jitter(alpha = 0.5) +
  facet_wrap(~ cluster)
```

Dot plot

```{r}
d <- hr_clust %>% 
  mutate(satisfaction_level = scales::rescale(satisfaction_level)) %>% 
  mutate(number_project = scales::rescale(number_project)) %>% 
  mutate(average_montly_hours = scales::rescale(average_montly_hours)) %>% 
  group_by(cluster) %>% 
  summarize(`Satisfaction` = mean(satisfaction_level),
            `Number of Projects` = mean(number_project),
            `Average Monthly Hours` = mean(average_montly_hours)) %>% 
  gather('variable', 'value', -cluster)

ggplot(d) +
  geom_point(aes(x = value, y = variable, color = cluster))
  

```


```{r}
hr1 <- hr %>% 
  filter(left == 1) %>% 
  select(time_spend_company, number_project)

km1 <- hr1 %>% 
  kmeans(centers = 3, nstart = 20)

hr1$cluster <- as.character(km1$cluster)

ggplot(hr1, aes(x = time_spend_company, y = number_project, color = cluster)) +
  geom_jitter(alpha = 0.5)
```



```{r}
library(randomForest)

hr_num <- hr %>% 
  select(left, satisfaction_level, last_evaluation, number_project, average_montly_hours, time_spend_company, Work_accident, promotion_last_5years) %>% 
  mutate(left = as.factor(left))

f <- randomForest(left ~ ., data = hr_num)

varImpPlot(f)
```

## Survival Model

```{r}
library(survival)
library(survminer)

hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  select(-sales)

x_surv <- Surv(time = hr$time_spend_company, event = hr$left)
f_surv <- survfit(x_surv ~ salary, data = hr)
ggsurvplot(f_surv, data = hr)

x_surv <- Surv(time = hr$time_spend_company, event = hr$left)
f_surv <- survfit(x_surv ~ salary, data = hr)
ggsurvplot(f_surv, data = hr)

# hr <- dummy_cols(hr, select_columns = c('salary'))
f_coxph <- coxph(x_surv ~ satisfaction_level + last_evaluation + 
                   I(number_project^2) +
                   number_project + average_montly_hours +
                   salary, data = hr)
ggforest(f_coxph)

d <- hr %>% 
  # filter(last_evaluation > 0.8, satisfaction_level > 0.8) %>% # super people
  filter(last_evaluation < 0.6, satisfaction_level < 0.4) %>% # losers
  mutate(satisfaction_level = as.numeric(scale(satisfaction_level))) %>% 
  mutate(last_evaluation = as.numeric(scale(last_evaluation))) %>% 
  mutate(number_project = as.numeric(scale(number_project))) %>% 
  mutate(average_montly_hours = as.numeric(scale(average_montly_hours))) %>% 
  mutate(salary = as.numeric(scale(salary)))
x_surv <- Surv(time = d$time_spend_company, event = d$left)
f_coxph <- coxph(x_surv ~ satisfaction_level + last_evaluation + 
                   I(number_project^2) + number_project + 
                   average_montly_hours + salary, 
                 data = d)
ggforest(f_coxph, data = d, main = 'Hazard ratio for Winners')


```

```{r}
x_surv <- Surv(time = hr$time_spend_company, 
               event = hr$promotion_last_5years)
f_surv <- survfit(x_surv ~ number_project, data = hr)
ggsurvplot(f_surv, data = hr)

f_coxph <- coxph(x_surv ~ satisfaction_level + last_evaluation + 
                   I(number_project^2) +
                   number_project + average_montly_hours, 
                 data = hr)
ggforest(f_coxph)
```

## Survival model plotting

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  select(-sales) %>% 
  mutate(num_project2 = number_project^2)

x_surv <- Surv(time = hr$time_spend_company, 
               event = hr$promotion_last_5years)
f_coxph <- coxph(x_surv ~ satisfaction_level + last_evaluation + 
                   num_project2 + number_project + average_montly_hours, 
                 data = hr)
```

```{r}
library(iml)
d <- hr %>% 
  select(satisfaction_level, last_evaluation, 
         num_project2, number_project, average_montly_hours, left) %>% 
  sample_frac(size = 0.2)
cox_pred_fun <- function(model, newdata) {
  exp(predict(model, newdata = newdata))
}
cox_pred <- Predictor$new(model = f_coxph, data = d, y = hr$left,
                          predict.fun = cox_pred_fun)

imp <- FeatureImp$new(cox_pred, loss = "ce")
plot(imp)
```




## Random forest to predict department

```{r}
rf_dept <- randomForest(department ~ satisfaction_level + 
                          last_evaluation + number_project + average_montly_hours +
                          time_spend_company + Work_accident, data = hr, ntree = 100)
varImpPlot(rf_dept)
```

## Residual salary

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) 
f_salary <- lm(salary ~ time_spend_company, data = hr)

residuals(f_salary) %>% qplot(bins = 100)

hr$salary_time_residual <- residuals(f_salary)


hr_num <- hr %>% 
  select(left, satisfaction_level, last_evaluation, number_project, average_montly_hours, time_spend_company, Work_accident, promotion_last_5years, salary_time_residual) %>% 
  mutate(left = as.factor(left))

library(randomForest)
hr_rf <- randomForest(left ~ ., data = hr_num, ntree = 100)
varImpPlot(hr_rf)
```

ROC for this model

```{r}
library(pROC)
d <- tibble(pred = predict(hr_rf, type = 'prob')[,2],
            obs = hr_num$left)
hr_rf_roc <- roc(d, obs, pred)
plot(hr_rf_roc, print.auc = T)
```

## Random Forest without Job Satisfaction

```{r}
library(randomForest)
hr_rf_nosat <- randomForest(left ~ ., 
                      data = hr_num %>% select(-satisfaction_level), 
                      ntree = 100)
varImpPlot(hr_rf_nosat)
d <- tibble(pred = predict(hr_rf_nosat, type = 'prob')[,2],
            obs = hr_num$left)
hr_rf_nosat_roc <- roc(d, obs, pred)
plot(hr_rf_nosat_roc, print.auc = T)
```


## Partial Dependence Plot
Need to install `iml`, need to install `devtools`

```{r}
library(iml)
library(randomForest)

hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  select(-sales) %>% 
  mutate(left = as.factor(left))

hr_rf <- randomForest(left ~ ., data = hr, ntree = 100)

rf_pred <- Predictor$new(hr_rf, type = 'prob', class = 2,
                        data = hr %>% select(-left), 
                        y = as.numeric(as.character(hr$left)))

rf_imp <- FeatureImp$new(rf_pred, loss = "f1")
plot(rf_imp)
```

```{r}
hr_rf_pdp <- FeatureEffect$new(rf_pred, method = 'pdp',
                                 feature = "satisfaction_level", 
                                 # center.at = 3,
                                 grid.size = 20)
plot(hr_rf_pdp) + theme_bw()
```

```{r}
hr_rf_pdp2 <- FeatureEffect$new(rf_pred, method = 'pdp', 
                            feature = c('satisfaction_level', 'last_evaluation'), grid.size = 10)
plot(hr_rf_pdp2) + 
  theme_bw() +
  scale_fill_viridis_c()
```

```{r}
hr_rf_ale <- FeatureEffect$new(rf_pred, method = 'ale',
                                 feature = c("satisfaction_level", "number_project"),
                                 # center.at = median(hr$satisfaction_level),
                                 grid.size = 100)
p1 <- plot(hr_rf_ale) + scale_fill_viridis_c() + theme_bw()

p2 <- ggplot(hr, aes(x = satisfaction_level, y = number_project)) +
  geom_point(position = position_jitter(height = 0.2), alpha = 0.2) +
  theme_minimal()

library(gridExtra)
grid.arrange(p1, p2)
```


## Shapley Values

```{r}
x_selected <- sample_n(hr_num, 1) %>% 
  select(-left)
hr_pred_shapley <- Shapley$new(rf_pred, x.interest = x_selected)
plot(hr_pred_shapley)
```


## Radar Chart

```{r}
library(fmsb)

hr <- read_csv('./kaggle_hr_analytics.csv')

d <- hr %>% 
  group_by(left) %>% 
  summarize(satisfaction_level = mean(satisfaction_level),
            last_evaluation = mean(last_evaluation),
            number_project = mean(number_project),
            average_montly_hours = mean(average_montly_hours),
            time_spend_company = mean(time_spend_company)) %>% 
  mutate(left = ifelse(left == 1, 'Left', 'Stayed'))

d_max <- hr %>% 
  select(one_of(c('satisfaction_level','last_evaluation','number_project','average_montly_hours','time_spend_company'))) %>% 
  summarize_all(max) %>% 
  mutate(left = 'max')
d_min <- hr %>% 
  select(one_of(c('satisfaction_level','last_evaluation','number_project','average_montly_hours','time_spend_company'))) %>% 
  summarize_all(min) %>% 
  mutate(left = 'min')

d <- bind_rows(d_max, d_min, d) %>% 
  as.data.frame

rownames(d) <- d$left
d$left <- NULL

radarchart(d, axistype = 0, 
           plwd = 3, plty = 1,
           cglcol="grey", cglty=1, axislabcol="grey", cglwd=0.8,
           vlcex=0.8)

```

## Radar chart data as a bar chart instead

```{r}
d <- hr %>% 
  select(one_of(c('left','satisfaction_level','last_evaluation','number_project','average_montly_hours','time_spend_company'))) %>% 
  mutate_all(scales::rescale) %>%
  group_by(left) %>% 
  summarize_all(mean) %>% 
  mutate(left = ifelse(left == 1, 'Left', 'Stayed')) %>% 
  gather(variable, value, -left)

ggplot(d) +
  geom_bar(stat = 'identity', position = 'dodge2',
           aes(x = left, y = value, fill = left)) +
  facet_wrap(~ variable, scales = 'free')

ggplot(d) +
  geom_point(size = 5, aes(x = value, y = variable, color = left)) +
  scale_y_discrete('') +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), 
        panel.grid.major.y = element_line(linetype = 2, color = 'grey40')) +
  geom_vline(xintercept = 0)
```

## Ridge plot

```{r}
library(ggridges)
d <- hr %>% 
  select(one_of(c('left','satisfaction_level','last_evaluation','number_project','average_montly_hours','time_spend_company'))) %>% 
  mutate_all(scales::rescale) %>%
  mutate(left = ifelse(left == 1, 'Left', 'Stayed')) %>% 
  gather(variable, value, -left)


ggplot(d, aes(x = value, y = variable, fill = left)) +
  geom_density_ridges(alpha = 0.6, color = 'white') +
  theme_ridges() + 
  theme(legend.position = "none")
```


## Decision tree model

```{r}
library(rpart)
library(rpart.plot)
library(pROC)

hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  select(-sales) %>% 
  mutate(left = as.factor(left))
```

```{r}
cv_set <- sample(1:10, size = nrow(hr), replace = T)

hr_test <- hr[cv_set == 1, ]
hr_train <- hr[cv_set != 1, ]

hr_dt <- rpart(left ~ satisfaction_level + last_evaluation + number_project + 
                 average_montly_hours + 
                 time_spend_company + Work_accident + 
                 promotion_last_5years + salary, 
               data = hr_train, 
               control = rpart.control(maxdepth = 3, cp = 0),
               model = T)
rpart.plot(hr_dt)
```

## K-fold cross validation

```{r}
kfolds <- 10
cv_folds <- sample(1:kfolds, size = nrow(hr), replace = T)

all_auc <- numeric(kfolds)
all_acc <- numeric(kfolds)
for(k in 1:kfolds) {
  print(k)
  hr_test <- hr[cv_folds == k, ]
  hr_train <- hr[cv_folds != k, ]
  
  hr_dt <- rpart(left ~ ., data = hr_train, control = rpart.control(maxdepth = 3))
  d <- tibble(pred = predict(hr_dt, newdata = hr_test)[,2],
              pred_class = predict(hr_dt, newdata = hr_test, type = 'class'),
              obs = hr_test$left)
  dt_roc_test <- roc(d, obs, pred)
  all_auc[k] <- auc(dt_roc_test)
  all_acc[k] <- sum(d$obs == d$pred_class) / nrow(hr_test)
}
mean(all_auc) # area under the curve
mean(all_acc) # accuracy
```


```{r}
library(caret)
d <- tibble(pred = ifelse(predict(hr_dt, newdata = hr_train) > 0.5, 'left', 'stayed') %>%  as.factor,
            obs = ifelse(hr_train$left == 1, 'left', 'stayed') %>% as.factor)
confusionMatrix(d$pred, d$obs)
```

```{r}
library(pROC)
d <- tibble(pred = predict(hr_dt, newdata = hr_train),
            obs = hr_train$left)
dt_roc_train <- roc(d, obs, pred)

d <- tibble(pred = predict(hr_dt, newdata = hr_test),
            obs = hr_test$left)
dt_roc_test <- roc(d, obs, pred, ci = T)

plot(dt_roc_train)

ciobj <- ci.se(dt_roc_test)
plot(ciobj, type = "shape", col = "#1c61b6AA") 
plot(dt_roc_test, print.auc = T,
     ci = T,col = 'firebrick4')
```


## Number of Projects and Satisfaction

```{r}
library(tidyverse)
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  select(-sales)

d <- hr %>% 
  mutate(left = ifelse(left == 1, 'Voluntary turnover', 'Stayed'))
ggplot(d, aes(x = number_project, y = satisfaction_level, color =  left)) +
  geom_jitter(alpha = 0.3) +
  scale_color_brewer('Turnover', type = 'qual', palette = 2) +
  scale_x_continuous('Number of Projects', breaks = 1:7) +
  scale_y_continuous('Job Satisfaction') +
  theme_minimal()
```

```{r}
lm_proj_satis <- lm(satisfaction_level ~ number_project + I(number_project^2), data = hr)
summary(lm_proj_satis)
```

```{r}
d <- hr %>% 
  mutate(left = ifelse(left == 1, 'Voluntary turnover', 'Stayed'))

x <- tibble(number_project = seq(2,7,length.out = 100))
x$pred <- predict(lm_proj_satis, newdata = x)
ggplot() +
  geom_jitter(alpha = 0.3, data = d, 
              aes(x = number_project, y = satisfaction_level, color =  left)) +
  geom_line(data = x, aes(x = number_project, y = pred)) +
  scale_color_brewer('Turnover', type = 'qual', palette = 2) +
  scale_x_continuous('Number of Projects', breaks = 1:7) +
  scale_y_continuous('Job Satisfaction') +
  coord_cartesian(xlim = c(2, 7), ylim = c(0, 1)) +
  theme_minimal()
```

## Random Forest as in report

```{r}
library(tidyverse)
library(randomForest)

hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  select(-sales) %>% 
  mutate(left = as.factor(left))
```

```{r}
hr_rf <- randomForest(left ~ ., data = hr,
                      importance = T, ntree = 100,
                      localImp = T)
varImpPlot(hr_rf)

importance(hr_rf)
```

## Random Forest Explainer
https://cran.rstudio.com/web/packages/randomForestExplainer/vignettes/randomForestExplainer.html

```{r}
library(randomForestExplainer)

min_depth <- min_depth_distribution(hr_rf)
plot_min_depth_distribution(min_depth)
```

```{r}
importance_frame <- measure_importance(hr_rf)
plot_multi_way_importance(importance_frame, size_measure = "no_of_nodes")

```



## Replicating the distplot

```{r}
p1 <- ggplot(hr, aes(x = satisfaction_level, y = ..density..)) + 
  geom_histogram(binwidth = 0.02, fill = '#b6c6df', color = 'white') + 
  geom_density(color = '#5a71a2') + 
  scale_x_continuous(limits = c(0,1)) +
  theme_classic()
p2 <- ggplot(hr, aes(x = last_evaluation, y = ..density..)) + 
  geom_histogram(binwidth = 0.02, fill = '#b6c6df', color = 'white') + 
  geom_density(color = '#5a71a2') + 
  scale_x_continuous(limits = c(0,1)) +
  theme_classic()
p3 <- ggplot(hr, aes(x = average_montly_hours, y = ..density..)) + 
  geom_histogram(binwidth = 5, fill = '#b6c6df', color = 'white') + 
  geom_density(color = '#5a71a2') +
  scale_x_continuous() +
  theme_classic()

library(gridExtra)

grid.arrange(p1, p2, p3, nrow = 1)

```

## Logistic Regression

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  select(-sales)

hrs <- hr %>% 
  mutate_at(c('satisfaction_level', 'last_evaluation',
              'number_project','average_montly_hours',
              'salary','time_spend_company', 'salary'), 
            .funs = function(x) as.numeric(scale(x)))

hr_log <- glm(left ~ ., data = hrs, family = 'binomial')
```

Coefficient plot for the logistic regression

```{r}
d1 <- coef(hr_log) %>% 
  exp %>% 
  as.data.frame %>% 
  as_tibble(rownames = 'variable')
colnames(d1) <- c('variable', 'coef')
d2 <- confint(hr_log) %>% 
  exp %>% 
  as.data.frame %>% 
  as_tibble(rownames = 'variable')
colnames(d2) <- c('variable', 'lower', 'upper')
hr_log_coef <- left_join(d1, d2)

ggplot(hr_log_coef) +
  geom_vline(xintercept = 1) +
  geom_point(aes(x = coef, y = variable)) +
  geom_segment(aes(x = lower, xend = upper, y = variable, yend = variable)) +
  theme_minimal() +
  theme()
```

Use fake data to explore the model results.

```{r}
d1 <- hrs %>% summarize_all(mean)
d2 <- expand.grid(promotion_last_5years = c(0,1), time_spend_company = seq(0,10,0.5))
d1 <- d1[rep(1, nrow(d2)),]
d1$promotion_last_5years <- d2$promotion_last_5years
d1$time_spend_company <- d2$time_spend_company

d1$pred <- predict(hr_log, newdata = d1, type = 'response')
d1$promotion_last_5years <- as.factor(d1$promotion_last_5years)

ggplot(d1, aes(x = time_spend_company, y = pred, color = promotion_last_5years)) +
  geom_line()
```

```{r}
library(iml)

log_pred <- function(model, newdata)  {
  results <- predict(model, newdata = newdata, type = 'response')
  return(results)
}
log_pred(hr_log, hrs) %>% head

p <- Predictor$new(hr_log, 
                   data = select(hrs, -left), 
                   y = hrs$left, predict.fun = log_pred,
                   class = 'classification')
imp <- FeatureImp$new(p, loss = "mse")
plot(imp)


```

```{r}
hr_log_pdp <- FeatureEffect$new(p, method = 'pdp',
                                 feature = "salary", 
                                 center.at = min(hrs$salary),
                                 grid.size = 20)
plot(hr_log_pdp) + theme_bw()
```


## Department Satisfaction histogram

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3)) %>% 
  mutate(department = sales) %>% 
  mutate(department = fct_infreq(department)) %>% 
  mutate(left = as.factor(left))

ggplot(hr, aes(x = satisfaction_level, fill = left, y = ..density..)) +
  geom_histogram(position = 'identity', alpha = 0.5, 
                 color = 'white', binwidth = 0.05) +
  scale_fill_brewer('Left',type = 'qual', palette = 2) +
  scale_x_continuous('Satisfaction Level') +
  facet_wrap(~ department) +
  theme_minimal()
```

```{r}
ggplot(hr, aes(x = number_project, fill = left, y = ..prop..)) +
  geom_bar(position = 'identity', alpha = 0.5, 
                 color = 'white') +
  scale_fill_brewer('Left',type = 'qual', palette = 2) +
  scale_x_continuous('Number of Projects') +
  facet_wrap(~ department, nrow = 2) +
  theme_minimal()

ggplot(hr, aes(x = number_project, y = ..prop..)) +
  geom_bar(position = 'identity', alpha = 0.5, 
                 color = 'white') +
  scale_x_continuous('Number of Projects') +
  facet_grid(department ~ left) +
  theme_minimal()
```


## An Isomap

```{r}
library(vegan)

d <- iris %>% 
  dummy_cols(select_columns = 'Species', 
             remove_first_dummy = T, ) %>% 
  select(-Species)

iris_dist <- vegdist(d)

ord <- isomap(iris_dist, epsilon = 0.18)
ord
pl <- plot(ord, main="isomap epsilon = 0.2")
```

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,
                            salary == "medium" ~ 2,
                            salary == "high" ~ 3))

d <- hr %>% 
  select(-left) %>% 
  select_if(is.numeric) %>% 
  sample_n(200)

```

## Job Satisfaction by Salary and Tenure

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = factor(salary, levels = c('low', 'medium', 'high'))) %>% 
  select(salary, satisfaction_level, tenure = time_spend_company)

hr %>% 
  group_by(salary, tenure) %>% 
  summarize(satisfaction_level = mean(satisfaction_level)) %>% 
  ggplot() +
  geom_line(aes(x = tenure, y = satisfaction_level, color = salary))

```

## Productivity and Mangement

```{r}
hr <- read_csv('./kaggle_hr_analytics.csv') %>% 
  mutate(salary = case_when(salary == "low" ~ 1,salary == "medium" ~ 2,salary == "high" ~ 3)) %>% 
  mutate(management = ifelse(sales == 'management', 'management', 'not management')) %>% 
  mutate(management = as.factor(management)) %>% 
  mutate(hourly_salary = salary / average_montly_hours) %>% 
  mutate(productivity = number_project / average_montly_hours)


ggplot(hr) +
  geom_histogram(aes(x = hourly_salary), bins = 100, color = 'white', size = 0.2)
ggplot(hr) +
  geom_histogram(aes(x = productivity), bins = 100, color = 'white', size = 0.2)

ggplot(hr, aes(x = hourly_salary, fill = management, y = ..density..)) +
  geom_histogram(position = 'identity', alpha = 0.5, 
                 color = 'white') +
  scale_fill_brewer('Management',type = 'qual', palette = 2) +
  scale_x_continuous('Salary per Monthly Hour') +
  theme_minimal()

ggplot(hr, aes(x = productivity, fill = management, y = ..density..)) +
  geom_histogram(position = 'identity', alpha = 0.5, 
                 color = 'white') +
  scale_fill_brewer('Management',type = 'qual', palette = 2) +
  scale_x_continuous('Productivity \n (number of projects / average hour worked)') +
  theme_minimal()

t.test(productivity ~ management, data = hr)

ggplot(hr, aes(x = productivity, y = hourly_salary, color = management)) +
  geom_point()

```

```{R Test tube}
a <- table(hr_new_oversample$satisfaction_level)
b <- table(hr_new_oversample$last_evaluation)

```

```{R Testing center}
m <- cmdscale(hr_new_oversample) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'provider') 

ggplot(m, aes(x = satisfaction_level)) +
  geom_point
```





















